---
title: "R Notebook"
output: html_notebook
---

```{r, setup}
library(magrittr)
library(dplyr)
library(ggplot2)
library(gapminder)
```

# Functional programming

```{r}
myFun <- function(x, f, ...){
 f(x, ...)
}

names <- c("Alice", "Bob", "Charlie")
toupper(names)
myFun(names, toupper)
myFun(names, rep, 2)
myFun(names, rep, each = 2)
```



# nested dataframes
## What is a data.frame?

## nested data frames are just lists of data frames 
```{r}
(by_country <- gapminder %>% 
  dplyr::group_by(country, continent) %>% 
  tidyr::nest())
```
```{r}
by_country$data[1]
```

## map functions work on lists
```{r}
map(by_country$data, ~{
  df <- .
  mean(df$lifeExp)
})
```
## using map inside mutate functions makes powerful combo
```{r}
by_country %>% 
  dplyr::mutate(avg_lifeExp = map(data, ~{mean(.$lifeExp)}))
```
## map_<type> returns non-list objects
```{r}
by_country %>% 
  dplyr::mutate(avg_lifeExp = map_dbl(data, ~{mean(.$lifeExp)})) 
```

```{r}
by_country %>% 
  dplyr::mutate(avg_lifeExp = map_dbl(data, ~{mean(.$lifeExp)})) %>% 
  tidyr::unnest(data) 
```


# nonstandard evaluation (ggplot, dplyr)

## iterate through all columns for plotting
```{r}
x.var <- names(diamonds)

map(x.var, ~{
  ggplot(diamonds, aes_string(., "price")) +
    geom_point()
})
```

## write function that fits lm to data 
input: data, lm, lm_name 

output: lm, broom::glance columns
```{r}

```

