---
title: "Functional Programming examples"
output: html_notebook
---
```{r}
library(magrittr)
```

```{r}
myFun <- function(x, fun, ...){
  fun(x, ...)
}
```

```{r}
myFun(c(0,20,101), mean)
myFun(c(0,20,101), median)
myFun(c(0,20,101), rep, 2)
```

```{r}
library(purrr)
rep2 <- function(x){rep(x, 2)}
funs <- c(mean, median, rep2)
dat <- c(0,20,101)
map2(list(dat), funs, ~{.y(.x)})
```

# What is a data.frame anyway?

```{r}
names <- c("Alice", "Bob", "Charlie")
age <- c(10, 20, 30)
pets <- c("cat", "dog", "fish")
(people <- data.frame(names, age, pets))
```
```{r}
people %>% str
```

```{r}
people %>% class
```
```{r}
people %>% typeof
```


## base data.frames coerce types so don't like this:
```{r}
pets <- list("cat", "dog", c("cat", "dog"))
(people <- data.frame(names, age, pets))
```
## Can force with I()
```{r}
(people <- data.frame(names, age, I(pets)))
```
```{r}
people %>% str
```

## Tibble won't coerce data types
```{r}
(people <- tibble::tibble(names, age, pets))
```

```{r}
str(people)
```

# data.frames are just named lists
by storing multiple things in a "cell", we can keep data organized in a single object


# operating on data subsets
```{r}
iris 
```

# Iris dataset:
```{r}
library(ggplot2)
iris %>% 
  ggplot(aes(Petal.Length, Sepal.Length)) +
    geom_point(aes(color = Species))
```

# fit lm to each species
## split-apply-combine
```{r}
by_species <- iris %>% 
  split(., .$Species)

by_species
```
```{r}
by_species_lm <- lapply(by_species, function(x){
  x %>% 
    lm(Sepal.Length~Petal.Length, data = .)
})

by_species_lm
```
# Now you've got two objects:
## What if you want to test multiple models? Manually type all this each time?
```{r}
by_species
by_species_lm
```



# Nested Data
Want to run the lm function on each row of `by_species$data`
```{r}
by_species <- iris %>% 
  dplyr::group_by(Species) %>% 
  tidyr::nest()

by_species
```


# Solutions for operating on multiple inputs:

## vectorized functions
```{r}
toupper(names)
```
## for loop
```{r}
upperNames <- rep("", length = length(names))

for (i in 1:length(names)) {
  upperNames[i] <- toupper(names[i])
}

upperNames
```

## Functional programming bypasses for-loop boilerplate
```{r}
myMap <- function(x, f){
  
  out <- rep("", length = length(x))
  
  for (i in 1:length(x)){
    out[i] <- f(x[i])
  }
  return(out)
}

names
myMap(names, toupper) 
myMap(names, tolower) 
```

## map function
`purrr` package provides a framework for many `map` functions
```{r}
library(purrr)
map(people$names, toupper) # operates on lists, returns list
map_chr(people$names, toupper) # operates on lists, returns character
```
# return to iris example:
```{r}
by_species
```
```{r}
by_species %>% 
  dplyr::mutate(fit = lm(Sepal.Length~Petal.Length, data = data))
```


```{r}
fit_model <- function(df){
  lm(Sepal.Length~Petal.Length, data = df)
}

(by_species %<>% 
  dplyr::mutate(fit = map(data, fit_model)))
```
```{r}
# use broom::glance to return tidy-data summary of each lm
(by_species %<>%
  dplyr::mutate(fit_summary = map(fit, broom::glance)))
```

```{r}
by_species$fit_summary 
```

# return specfic summary stats
```{r}
by_species %<>% 
  dplyr::mutate(rsq = map_dbl(fit_summary, ~{.$r.squared}))
```

# unnest to get tidy data version of each lm
```{r}
by_species %>% 
  tidyr::unnest(fit_summary)
```

```{r, fig.asp = 1}
by_species %>% 
  tidyr::unnest(data) %>% 
  dplyr::mutate(Species = paste0(Species, " (", round(rsq, 2), ")")) %>% 
  ggplot(aes(Petal.Length, Sepal.Length)) +
    geom_point(aes(color = Species)) +
    geom_smooth(aes(group = Species, color = Species), method = "lm", se = F) +
    xlim(0,8) +
    ylim(4,8) +
    theme(legend.position = "bottom")
```


```{r}
library(ggplot2)
gapminder <- gapminder::gapminder
gapminder
```

# functionals for plotting many vars
```{r}
library(ggplot2)
y.vars <- names(gapminder)[4:6]
y.vars

map(y.vars, ~{
  y.var <- .
  gapminder %>% 
    ggplot(aes_string("year", y.var)) +
      geom_line(aes(color = country, group = country)) +
      theme(legend.position = "none") + 
      facet_wrap(~continent, scales = "free_y") 
})
```

```{r}
print("TEST")
```




```{r}
by_continent <- gapminder %>% 
  dplyr::group_by(continent) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(plot = map2(continent, data, ~{
    .y %>% 
      ggplot(aes(year, lifeExp)) +
        geom_line(aes(color = country)) +
        ggtitle(.x) +
        theme(legend.position = "none")
  }))

by_continent$plot  
```

```{r}
mod <- function(data){
  lm(lifeExp~year, data = data)
}
```

```{r}
by_country <- gapminder %>% 
  dplyr::group_by(country, continent) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(lm_year_lifeExp = map(data, mod)) %>% 
  dplyr::mutate(lm_year_lifeExp_glance = map(lm_year_lifeExp, broom::glance)) %>% 
  dplyr::mutate(lm_year_lifeExp_goodFit = map_lgl(lm_year_lifeExp_glance, ~{.x$r.squared > 0.75}))

by_country %>% 
  ggplot(aes(continent)) +
    geom_bar(aes(fill = lm_year_lifeExp_goodFit))
```

# Advanced use:
```{r}
# make function that will fit linear model and add nested residuals
lm_fit <- function(df, lm, lm_name){
 
  # create column names 
  glance_name <- paste0(lm_name, "_glance") 
  fit_name <- paste0(lm_name, "_goodFit") 
  resid_name <- paste0(lm_name, "_resid") 
  
  # aggregate column names (creation & aggregation can be done in 1 step with a map function, but we'll keep it simple) 
  colNames <- c(glance_name, fit_name, resid_name)
  
  
  # use non-standard evaluation to create columns for each lm 
  df_out <- df %>% 
    dplyr::mutate(!! lm_name := map(data, lm)) %>% 
    dplyr::mutate(!! glance_name := map(.[[lm_name]], broom::glance)) %>% 
    dplyr::mutate(!! fit_name := map_lgl(.[[glance_name]], ~{.x$r.squared > 0.75})) %>% 
    dplyr::mutate(!! resid_name := map2(data, .[[lm_name]], modelr::add_residuals)) %>% 
    dplyr::select(colNames) # for this example, we'll only return the columns we create
  
  return(df_out)
}

# make functions for each model:
mod <- function(df){
  lm(lifeExp~year, data = df)
}

mod2 <- function(df){
  lm(lifeExp~year+gdpPercap, data = df)
}

# make data.frame containing & describing each model:
mods <- tibble::tribble(~"mod", ~"name",
                mod, "lm_lifeExp_year",
                mod2, "lm_lifeExp_year.gdp")

# get data in nested format:
by_country <- gapminder %>% 
  dplyr::group_by(continent, country) %>% 
  tidyr::nest()


# run each lm on nested data
lms <- map2(mods$mod, mods$name, ~{
  lm_fit(by_country, .x, .y)
})

# add lm information to nested data
by_country %<>% dplyr::bind_cols(lms)
```

# compare lms
One advantage of having the `mods` data.frame is that you can split your data on it later
```{r}
map(mods$name, ~{
  
  colname <- paste0(., "_resid")
  
  by_country %>% 
    tidyr::unnest_(colname) %>% 
    ggplot(aes(year, resid)) +
    geom_line(aes(group = country), alpha = 1/10) +
    geom_smooth(method = "gam") +
    ggtitle(colname)
})
```


```{r}
# test passing map function as string to mutate_


dat %>% 
  dplyr::mutate_(mapTest = glue::glue("map(data, {lm_fun})", column = "out", lm_fun = "mod"))


dat %>% 
  dplyr::mutate(!!colNames[1] := map(data, mod)) %>% 
  dplyr::mutate(!!colNames[2] := map(colNames[1], broom::glance))
                  
                  
```


```{r}
library(gapminder)
library(dplyr)
library(purrr)

colNames <- c("my_lm", "my_lm_glance")
col1 <- colNames[1]
col2 <- colNames[2]

mod <- function(data){
  lm(lifeExp~year, data = data)
}

gapminder %>%
  dplyr::group_by(country, continent) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(!! col1 := map(data, mod)) %>% 
  dplyr::mutate(!! col2 := map(.[[col1]], broom::glance)) 
  
```

```{r}
l1 <- list(list(a = 1L), list(a = NULL, b = 2L), list(b = 3L))
col <- "a"
l1 %>% map(col, .default = "???")
l1 %>% map_int("b", .default = NA)
l1 %>% map_int(2, .default = NA)

```




```{r}
lm_fit <- function(df, lm, lm_name){
  
  colNames <- map2(lm_name, c("", "_glance", "_goodFit", "_resid"), ~{glue::glue("{name}{colType}", name = .x, colType = .y)}) %>% 
    unlist
  
   
  df_out <- df %>% 
    dplyr::mutate(!! colNames[1] := map(data, lm)) %>% 
    dplyr::mutate(!! colNames[2] := map(.[[colNames[1]]], broom::glance)) %>% 
    dplyr::mutate(!! colNames[3] := map_lgl(.[[colNames[2]]], ~{.x$r.squared > 0.75})) %>% 
    dplyr::mutate(!! colNames[4] := map2(data, .[[colNames[1]]], modelr::add_residuals)) %>% 
    dplyr::select(colNames)
  
  return(df_out)
}
```

# more lm examples:
```{r}
lm(lifeExp ~ year, gapminder) %>% summary
lm(lifeExp ~ year + gdpPercap, gapminder) %>% summary
```

```{r}
fit_lm <- function(df, fun){
  lm(fun, data = df)
}

funs <- c(lifeExp ~ year, lifeExp ~ year + gdpPercap)
map(funs, ~{fit_lm(gapminder, .) %>% summary})
```
